---
title: "Using tidyhla for HLA Typing Data Processing"
author: "Geeta Karadkhele and Chris Larsen"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using tidyhla for HLA Typing Data Processing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

### Introduction

The tidyhla package simplifies the preprocessing, cleaning, and analysis of HLA typing data. It provides functions to standardize allele formats, evaluate donor-recipient mismatches, and clean messy data efficiently. This vignette will guide you through key functionalities with real-world examples.

```{r}
# Install devtools if not already installed
install.packages("devtools")

# Install tidyhla from GitHub
devtools::install_github("LarsenLab/tidyhla")
```


```{r setup}
library(tidyhla)
```


#### Cleaning HLA Typing Data

##### Example: HLA Class I Cleaning
##### Text: Use clean_hla_class1() to clean and standardize Class I HLA typing data. 

```{r}
# Example data
hla_data <- data.frame(
  allele1 = c("A*01:01N", "A*02:01", "A*03:02G"),
  allele2 = c("A*01:02P", "A*02:03", "A*03:04")
)

# Clean Class I data
cleaned_class1 <- clean_hla_class1(hla_data, var_1 = allele1, var_2 = allele2)
cleaned_class1
```

#### Evaluating Mismatches
##### Example data 
##### Text: Use eval_mism() to evaluate mismatches between donor and recipient alleles.

```{r}
mismatch_data <- data.frame(
  donor_allele1 = c("A*01:01", "A*02:01"),
  donor_allele2 = c("A*01:02", "A*02:03"),
  recipient_allele1 = c("A*01:01", "A*03:04"),
  recipient_allele2 = c("A*01:02", "A*03:05")
)

# Evaluate mismatches
mismatches <- eval_mism(
  mismatch_data,
  don_1 = donor_allele1,
  don_2 = donor_allele2,
  recip_1 = recipient_allele1,
  recip_2 = recipient_allele2
)
mismatches
```


```{r}
sessionInfo()
```

